They're Clutching up! Team Momentum in Round-Based Esports

### 2022-08-14 paper2/paper2.tex
paper 2

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={They're Clutching up! Team Momentum in Round-Based Esports},
  pdfauthor={Tony ElHabr},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{They're Clutching up! Team Momentum in Round-Based Esports}
\author{
Tony ElHabr\\\\Georgia Institute of
Technology\\\\\href{mailto:anthonyelhabr@gmail.com}{anthonyelhabr@gmail.com}}
\date{}
\begin{document}
\maketitle
\begin{abstract}
My research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD).

@@

Second, I examine post-streak round win probability. I find that teams
perform significantly worse than expected after streaks of 2, 3, and 4
wins when series end up going to 9, 10, or 11 (maximum) rounds, even
after accounting for the ``hot-hand'' phenomenon.

Third, I compare win percentages in round one versus all other rounds,
hypothesizing that there may be some advantage on either side when there
is no prior information about how the opponent intends to play a given
map in either game. I find only one instance for which there seems to be
a significant defensive advantage in round one.

Finally, I evaluate behavior when teams have two rounds left to win the
series, observing a peak in COD offensive win percentages in the 4-4
state, and no such oddity in Valorant.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, enhanced, breakable, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

Call of Duty (CoD), first released in 2003, is one of the most popular
first-person shooter (FPS) video game franchises of all-time. Tthe most
popular mode in the competitive scene is ``Search and Destroy'' (SnD),
which bears resemblance to ``Bomb Defusal'' in Counter-Strike and
``Plant/Defuse'' in Valorant, two other FPS games played in professional
leagues. SnD is one-sided game mode in which one team, the offensive
side, tries to destroy one of two designated bomb sites on the map.

In professional CoD, a team must win six rounds of SnD to win the
match.\footnote{A maximum of 11 even rounds can be played. There is no
  ``sudden death'' or ``win by two'' rule like there are for SnD
  equivalent in professional Counter-Strike and Valorant matches.} A
round can end in one of four ways:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  One team eliminates all members of the other team prior to a bomb
  plant. (Eliminating team wins.)
\item
  The offensive team eliminates all members of the defensive team after
  a bomb plant.\footnote{Merchanics regarding the bomb:} (Offense wins.)
\item
  The defensive team defuses the bomb after a bomb plant.\footnote{Often
    the defensive team will try to eliminate all team members prior to
    making the defuse, but in some cases, they may try to ``ninja''
    defuse.} (Defense wins.)
\item
  The offensive team does not make a plant by the time the round timer
  ends. (Defense wins.)
\end{enumerate}

\begin{itemize}
\tightlist
\item
  The bomb can be picked up by any member of the offensive team.
\item
  The bomb carrier is not obstructed at all by carrying the bomb
  (i.e.~movement is the same, weapon usage is the same).
\item
  The defense does not get any visual indication for who is carrying the
  bomb.
\item
  A bomb plant takes five seconds. The timer resets if the player stops
  planting site prior to completing it.
\item
  A bomb defuse takes seven seconds. The timer resets if the player
  ``drops'' the bomb.
\item
  The bomb takes 45 seconds to defuse after being planted.
\end{itemize}

@@

We adopt the terminology ``series'' to refer to what CoD SnD players
typically call a ``match'', so as to emulate the terminology of playoff
series in professional leagues like the National Basketball Association,
National Hockey League, and Major League Baseball. A ``game'' or a
``match'' in such leagues is analogous to a ``round'' of CoD SnD.

@@

CoD has roughly gone through three eras of professional gaming: (1)
Major League Gaming (MLG) tournaments prior to 2016; (2) the CoD World
League (CWL), initiated in 2016; and (3) the 12-franchise CoD League
(CDL), running since 2020 and completing three year-long ``seasons''
completed as of August 2022.\footnote{CoD is fairly unique compared to
  other esports in that it runs on an annual lifecycle (released coming
  in the late fall), where a new game is published every year under the
  same title. Each new game bears resemblance to past ones, often
  introducing relatively small variations (``improvements'') to
  graphics, game modes, and other facets of gameplay. During the CDL
  era, the games released have been Modern Warfare (2020), Cold War
  (2021) and Vanguard (2022).} The data set consists of all SnD matches
played in tournaments and qualifiers during the CDL era, totaling 15,584
rounds across 1,704 series. Data was collected in spreadsheets by
community member ``IOUTurtle''.\footnote{Data:
  https://linktr.ee/CDLArchive. Author: https://twitter.com/IOUTurtle}

The empirical offensive round win percentage across all rounds is
47.8\%.\footnote{Offensive round win percentage has been nearly constant
  across the three games during the CDL era: MW (2020) 47.2\%, Cold War
  (2021) 47.9\%, Vanguard (2022) 48.1\%} To contextualize our subsequent
discussion of series length and momentum, table 1 shows round win
percentages by series ``state'' (i.e.~the number of round wins by each
team prior to an upcoming round). Offensive round win rate is not quite
constant, although never veers more than 10\% from this global average.

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{rllllll}
\toprule
 & \multicolumn{6}{c}{Offensive team's round wins} \\ 
\cmidrule(r){2-7}
Defensive team's round wins & 0 & 1 & 2 & 3 & 4 & 5 \\ 
\midrule
0 & 47.8\%
(852) & 46.6\%
(408) & 43.1\%
(216) & 43.5\%
(115) & 43.3\%
(67) & 40.5\%
(37) \\ 
1 & 48.6\%
(444) & 49.3\%
(418) & 51.5\%
(309) & 43.4\%
(205) & 43.3\%
(120) & 39.4\%
(99) \\ 
2 & 52.8\%
(218) & 48.9\%
(305) & 48.9\%
(315) & 46.6\%
(262) & 48.7\%
(189) & 42.1\%
(133) \\ 
3 & 54.5\%
(123) & 46.0\%
(200) & 49.6\%
(250) & 45.6\%
(248) & 44.4\%
(214) & 44.8\%
(174) \\ 
4 & 56.9\%
(65) & 54.5\%
(145) & 47.2\%
(193) & 44.7\%
(228) & 55.2\%
(221) & 50.5\%
(208) \\ 
5 & 47.4\%
(38) & 49.4\%
(83) & 47.1\%
(136) & 50.9\%
(175) & 45.2\%
(177) & 46.0\%
(202) \\ 
\bottomrule
\end{longtable}

@@

There have been only a handful of studies of the distribution of games
played in a series, most of which assume a constant probability \(p\) of
a given team winning a game in the series, regardless of the series
state. (Mosteller 1952) observed that the American League had dominated
the National League in Major League Baseball's (MLB) World Series
matchups, implying that matchups should not modeled with \$p=0.5\$.
Mosteller proposed three approaches for identifying the optimal constant
probability value of the stronger team in the World Series, finding
\(p \approx 0.65\). in each case: (1) a ``method of moments'' approach
in which one solves for \(p\). from the empirical average number of
games won by the loser of the series; (2) maximizing the likelihood that
the sample would have been drawn from a population in which the
probability of a team winning a game is constant across the series, and
(3) minimizing the chi-square goodness of fit estimate for \(p\).

(Chance 2020) re-examines the constant probability notion in Major
League Baseball's World Series (1923--2018), the National Basketball
Association's Finals (1951--2018), and the National Hockey League's
Stanley Cup (1939--2018). Chance applies finds strong evidence against
the null hypothesis of \(p = 0.5\) in the MLB and NHL championship
series when applying Mosteller's first and second methods.\footnote{Chance
  goes on to outline a conditional probability framework (likelihood of
  winning a game given the series state) which can exactly explain the
  distribution of the number of games played.} Chance's work is closely
related to ours and, in fact, provides a guide for the first part of our
investigation.

@@

Studying both player streaks and team streaks in basketball, in both
observational and controlled settings. (Gilovich, Vallone, and Tversky
1985) do not find evidence for the hot hand phenomenon. Recently ,
(Miller and Sanjurjo 2018) refute (Gilovich, Vallone, and Tversky
1985)'s conclusions, finding mathematical evidence that seems to support
negative recency. Specifically, they find that a ``bias exists in a
common measure of the conditional dependence of present outcomes on
streaks of past outcomes in sequential data'' (streak selection bias)
that imply that, under i.i.d. conditions, ``the proportion of successes
among the outcomes that immediately follow a streak of consecutive
successes is expected to be strictly less than the underlying
(conditional) probability of success''. We agree with Miller's findings,
accounting for the streak selection bias in our study of momentum.

@@

The general formula for the probability \(P_E(i)\) that a best-of-\(s\)
series lasts \(i\) rounds given constant probability \(p\) of one
team\footnote{If \(p > 0.5\), then we might say that this team is the
  better team (presumably known in hindsight)} winning each round is

@@

\begin{equation}\protect\hypertarget{eq-series-length}{}{
P_E(i) = \frac{(i - 1)!}{s_1!(i - s_2)!}(p^{s_2}q^{s_2^{'}} + p^{s_2^{'}}q^{s_2}).
}\label{eq-series-length}\end{equation}

Table 1 shows the expected proportion for \(s = 11\) and the naive
assumption that \(p = 0.5\), along with the observed proportion
\(P_{O}(i)\) for CoD SnD.

\begin{table}

\caption{\label{tbl-panel}: The probabilities that a best-of-11 series
lasts \(i\) rounds (\(P(i)\), where \(i \in R = [6, 7, 8, 9, 10, 11]\))
under the assumption that each team has a 50\% probability (\(p = 0.5\))
of winning each game \{\#tbl-prob-series-lasting-i-rounds\} along with
the observed frequencies of series lasting \(i\) rounds, expressed as a
percentage \(P_O(i)\) and as a count
(\(N_O\))}\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-panel-1}}

{\centering

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{rrrr}
\toprule
&  & \multicolumn{2}{c}{Observed} \\ 
\cmidrule(lr){3-4}
Series lasts $i$ rounds & $P_E(i)$ & $P_O(i)$ & $N_O(i)$ \\ 
\midrule
6 & $3.1\%$ & $4.7\%$ & 40 \\ 
7 & $9.4\%$ & $11.9\%$ & 101 \\ 
8 & $16.4\%$ & $16.5\%$ & 141 \\ 
9 & $21.9\%$ & $21.7\%$ & 185 \\ 
10 & $24.6\%$ & $21.5\%$ & 183 \\ 
11 & $24.6\%$ & $23.7\%$ & 202 \\ 
\bottomrule
\end{longtable}

}

\end{minipage}%

\end{table}

@@

\[
\chi^2 = \sum^{11}_{i=6} \frac{(P_O(i) - P_E(i))^2}{P_E(i)}, i \in R = [6, 7, 8, 9, 10, 11]
\]

as 16.0 (p-value of 0.0068), we can comfortably reject the constant
probability null hypothesis, event at a confidence level of
\(\alpha = 0.01\).

Table~\ref{tbl-mosteller-methods} shows the alternate values for the
constant round win probability \(p\) for winning a given round in a CoD
SnD that we find when applying the three methods suggested by (Mosteller
1952). Each is approximately or equal to 0.575, and each results in a
\(chi^2\) value for which we cannot reject constant probability null
hypothesis.

\hypertarget{tbl-mosteller-methods}{}
\begin{longtable}[]{@{}lrr@{}}
\caption{\label{tbl-mosteller-methods}Alternate estimates of the
constant probability \(p\) for winning a given round in a CoD SnD,
applying the three methods suggested by (Mosteller 1952), in addition to
the naive \(p = 0.5\).}\tabularnewline
\toprule()
Method & \(p\) & \(\chi^2\) (p-value) \\
\midrule()
\endfirsthead
\toprule()
Method & \(p\) & \(\chi^2\) (p-value) \\
\midrule()
\endhead
0. Naive & 0.5000 & 16.0 (\textless=0.01) \\
1. Method of moments & 0.5725 & 3.6 (0.6) \\
2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
3. Minimum \(\chi^2\) & 0.5775 & 3.5 (0.62) \\
\bottomrule()
\end{longtable}

Table~\ref{tbl-expected-series-lengths-alternative-ps} shows the new
\(P_E(i)\) when re-applying Equation~\ref{eq-series-length} for each new
\(p\). We observe that \(P_E(i)\) is notably smaller for \(p_{1,2,3}\)
when \(i \in [10, 11]\) and higher for \(i \in [6, 7]\), more closely
matching \(P_O(i)\).

>>>>

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{They're Clutching up! Team Momentum in Round-Based Esports}

\author{
    Tony ElHabr
   \\
     \\
   \\
  \texttt{\href{mailto:anthonyelhabr@gmail.com}{\nolinkurl{anthonyelhabr@gmail.com}}} \\
  }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\begin{document}
\maketitle

\begin{abstract}
My research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD).

@@

Second, I examine post-streak round win probability. I find that teams
perform significantly worse than expected after streaks of 2, 3, and 4
wins when series end up going to 9, 10, or 11 (maximum) rounds, even
after accounting for the ``hot-hand'' phenomenon.
\end{abstract}

@@

Call of Duty (CoD), first released in 2003, is one of the most popular
first-person shooter (FPS) video game franchises of all-time. The most
popular mode in the competitive scene is ``Search and Destroy'' (SnD),
which bears resemblance to ``Bomb Defusal'' in Counter-Strike and
``Plant/Defuse'' in Valorant, two other FPS games played in professional
leagues. SnD is one-sided game mode in which one team, the offensive
side, tries to destroy one of two designated bomb sites on the map.

In professional CoD, a team must win six rounds of SnD to win the
match.\footnote{A maximum of 11 even rounds can be played. There is no
  ``sudden death'' or ``win by two'' rule like there are for SnD
  equivalent in professional Counter-Strike and Valorant matches.} A
round can end in one of fmy ways:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  One team eliminates all members of the other team prior to a bomb
  plant. (Eliminating team wins.)
\item
  The offensive team eliminates all members of the defensive team after
  a bomb plant.\footnote{\begin{itemize}
    \tightlist
    \item
      The bomb can be picked up by any member of the offensive team.
    \item
      The bomb carrier is not obstructed at all by carrying the bomb
      (i.e.~movement is the same, weapon usage is the same).
    \item
      The defense does not get any visual indication for who is carrying
      the bomb.
    \item
      A bomb plant takes five seconds. The timer resets if the player
      stops planting site prior to completing it.
    \item
      A bomb defuse takes seven seconds. The timer resets if the player
      ``drops'' the bomb.
    \item
      The bomb takes 45 seconds to defuse after being planted.
    \end{itemize}} (Offense wins.)
\item
  The defensive team defuses the bomb after a bomb plant.\footnote{Often
    the defensive team will try to eliminate all team members prior to
    making the defuse, but in some cases, they may try to ``ninja''
    defuse.} (Defense wins.)
\item
  The offensive team does not make a plant by the time the round timer
  ends. (Defense wins.)
\end{enumerate}

@@

I adopt the terminology ``series'' to refer to what CoD SnD players
typically call a ``match'', so as to emulate the terminology of playoff
series in professional leagues like the National Basketball Association,
National Hockey League, and Major League Baseball. A ``game'' or a
``match'' in such leagues is analogous to a ``round'' of CoD SnD.

@@

CoD has roughly gone through three eras of professional gaming: (1)
Major League Gaming (MLG) tmynaments prior to 2016; (2) the CoD World
League (CWL), initiated in 2016; and (3) the 12-franchise CoD League
(CDL), running since 2020 and completing three year-long ``seasons''
completed as of August 2022.\footnote{CoD is fairly unique compared to
  other esports in that it runs on an annual lifecycle (released coming
  in the late fall), where a new game is published every year under the
  same title. Each new game bears resemblance to past ones, often
  introducing relatively small variations (``improvements'') to
  graphics, game modes, and other facets of gameplay. During the CDL
  era, the games released have been Modern Warfare (2020), Cold War
  (2021) and Vanguard (2022).} The data set consists of all SnD matches
played in tmynaments and qualifiers during the CDL era, totaling 7,792
rounds across 852 series. Data was collected in spreadsheets by
community member ``IOUTurtle''.\footnote{Data:
  \url{https://linktr.ee/CDLArchive}. Author:
  \url{https://twitter.com/IOUTurtle}}

The empirical offensive round win percentage across all rounds is
47.8\%.\footnote{Offensive round win percentage has been nearly constant
  across the three games during the CDL era: 1. 47.2\% in MW (2020) 2.
  47.9\% in Cold War (2021) 3. 48.1\% in Vanguard (2022)} To
contextualize my subsequent discussion of series length and momentum,
table \ref{tbl:cod-o-win-prop-by-series-state} shows round win
percentages by series ``state'' (i.e.~the number of round wins by each
team prior to an upcoming round). Offensive round win rate is not quite
constant, although never veers more than 10\% from this global average.

\begin{longtable}{crrrrrr}
  \caption{Offensive round win rates for the upcoming round, given both the offensive and defensive team's prior number of round wins}\label{tbl:cod-o-win-prop-by-series-state} \\
  \toprule
   & \multicolumn{6}{c}{Offense round wins} \\ 
  \cmidrule(lr){2-7}
  Defense round wins & 0 & 1 & 2 & 3 & 4 & 5 \\ 
    \midrule
    0 & 47.8\%
    (852) & 46.6\%
    (408) & 43.1\%
    (216) & 43.5\%
    (115) & 43.3\%
    (67) & 40.5\%
    (37) \\ 
    1 & 48.6\%
    (444) & 49.3\%
    (418) & 51.5\%
    (309) & 43.4\%
    (205) & 43.3\%
    (120) & 39.4\%
    (99) \\ 
    2 & 52.8\%
    (218) & 48.9\%
    (305) & 48.9\%
    (315) & 46.6\%
    (262) & 48.7\%
    (189) & 42.1\%
    (133) \\ 
    3 & 54.5\%
    (123) & 46.0\%
    (200) & 49.6\%
    (250) & 45.6\%
    (248) & 44.4\%
    (214) & 44.8\%
    (174) \\ 
    4 & 56.9\%
    (65) & 54.5\%
    (145) & 47.2\%
    (193) & 44.7\%
    (228) & 55.2\%
    (221) & 50.5\%
    (208) \\ 
    5 & 47.4\%
    (38) & 49.4\%
    (83) & 47.1\%
    (136) & 50.9\%
    (175) & 45.2\%
    (177) & 46.0\%
    (202) \\ 
  \bottomrule
\end{longtable}

@@

There have been only a handful of studies of the distribution of games
played in a series, most of which assume a constant probability \(p\) of
a given team winning a game in the series, regardless of the series
state. (Mosteller 1952) observed that the American League had dominated
the National League in Major League Baseball's (MLB) World Series
matchups, implying that matchups should not modeled with \(p = 0.5\).
Mosteller proposed three approaches for identifying the optimal constant
probability value of the stronger team in the World Series, finding
\(p \approx 0.65\). in each case: (1) a ``method of moments'' approach
in which one solves for \(p\). from the empirical average number of
games won by the loser of the series; (2) maximizing the likelihood that
the sample would have been drawn from a population in which the
probability of a team winning a game is constant across the series, and
(3) minimizing the chi-square goodness of fit estimate for \(p\).

(Chance 2020) re-examines the constant probability notion in Major
League Baseball's World Series (1923--2018), the National Basketball
Association's Finals (1951--2018), and the National Hockey League's
Stanley Cup (1939--2018). Chance applies finds strong evidence against
the null hypothesis of \(p = 0.5\) in the MLB and NHL championship
series when applying Mosteller's first and second methods.\footnote{Chance
  goes on to outline a conditional probability framework (likelihood of
  winning a game given the series state) which can exactly explain the
  distribution of the number of games played.} Chance's work is closely
related to mys and, in fact, provides a guide for the first part of my
investigation.

@@

Studying both player streaks and team streaks in basketball, in both
observational and controlled settings. (Gilovich, Vallone, and Tversky
1985) do not find evidence for the hot hand phenomenon. Recently ,
(Miller and Sanjurjo 2018) refute (Gilovich, Vallone, and Tversky
1985)'s conclusions, finding mathematical evidence that seems to support
negative recency. Specifically, they find that a ``bias exists in a
common measure of the conditional dependence of present outcomes on
streaks of past outcomes in sequential data'' (streak selection bias)
that imply that, under i.i.d. conditions, ``the proportion of successes
among the outcomes that immediately follow a streak of consecutive
successes is expected to be strictly less than the underlying
(conditional) probability of success''. I agree with Miller's findings,
accounting for the streak selection bias in my study of momentum.

@@

The general formula for the probability \(P_E(i)\) that a best-of-\(s\)
series lasts \(i\) rounds given constant probability \(p\) of one
team\footnote{If \(p > 0.5\), then I might say that this team is the
  better team (known omnisciently).} winning each round is

@@

\begin{equation}\protect\hypertarget{eq-series-length}{}{
P_E(i) = \frac{(i - 1)!}{s_1!(i - s_2)!}(p^{s_2}q^{s_2^{'}} + p^{s_2^{'}}q^{s_2}).
}\label{eq:series-length}\end{equation}

Table \ref{tbl:cod-prob-series-lasting-i-rounds} shows the expected
proportion for \(s = 11\) and the naive assumption that \(p = 0.5\),
along with the observed frequencies observed in CoD SnD, as both a count
\(N_O(i)\) (of series lasting \(i\) rounds) and a proportion \(P_O(i)\)
of all series.

\begin{longtable}{rrrr}
  \caption{The probabilities that a best-of-11 series lasts $i$ rounds ($Pr(i)$, where $i = {6, 7, 8, 9, 10, 11}$) under the assumption that each team has a 50% probability ($p = 0.5$) of winning each round.}\label{tbl:cod-prob-series-lasting-i-rounds} \\
  \toprule
  &  & \multicolumn{2}{c}{Observed} \\ 
  \cmidrule(lr){3-4}
  Series lasts $i$ rounds & $P_E(i)$ & $P_O(i)$ & $N_O(i)$ \\ 
  \midrule
  6 & $3.1\%$ & $4.7\%$ & 40 \\ 
  7 & $9.4\%$ & $11.9\%$ & 101 \\ 
  8 & $16.4\%$ & $16.5\%$ & 141 \\ 
  9 & $21.9\%$ & $21.7\%$ & 185 \\ 
  10 & $24.6\%$ & $21.5\%$ & 183 \\ 
  11 & $24.6\%$ & $23.7\%$ & 202 \\ 
  \bottomrule
\end{longtable}

The probabilities that a best-of-11 series lasts \(i\) rounds (\(P(i)\),
where \(i \in R = [6, 7, 8, 9, 10, 11]\)) under the assumption that each
team has a 50\% probability (\(p = 0.5\)) of winning each game
\{\#tbl-prob-series-lasting-i-rounds\} along with the observed
frequencies of series lasting \(i\) rounds, expressed as a percentage
\(P_O(i)\) and as a count (\(N_O\))

@@

\begin{equation}\protect\hypertarget{eq-series-length}{}{
\chi^2 = \sum^{11}_{i=6} \frac{(P_O(i) - P_E(i))^2}{P_E(i)}, i \in R = [6, 7, 8, 9, 10, 11].
}\label{eq:chi-squ-test}\end{equation}

as 16.0 (p-value of 0.0068), I can comfortably reject the constant
probability null hypothesis, event at a confidence level of
\(\alpha = 0.01\).

Table \ref{tbl:mosteller-methods-results} shows the alternate values for
the constant round win probability \(p\) for winning a given round in a
CoD SnD that I find when applying the three methods suggested by
Mosteller (1952). Each is approximately or equal to 0.575, and, when
applying Equation \ref{eq:chi-squ-test}, each results in a \(\chi^2\)
value for which I cannot reject the constant probability null
hypothesis.

\begin{longtable}[]{@{}lrr@{}}
  \caption{Alternate estimates of the constant probability \(p\) for winning a given round in a CoD SnD, applying the three methods suggested by Mosteller (1952), in addition to the naive \(p = 0.5\).}\label{tbl:mosteller-methods-results} \\
  \toprule()
    Method & \(p\) & \(\chi^2\) (p-value) \\
    \midrule()
    \endfirsthead
    \toprule()
    Method & \(p\) & \(\chi^2\) (p-value) \\
    \midrule()
  \endhead
  0. Naive & 0.5000 & 16.0 (\textless=0.01) \\
  1. Method of moments & 0.5725 & 3.6 (0.6) \\
  2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
  3. Minimum \(\chi^2\) & 0.5775 & 3.5 (0.62) \\
  \bottomrule()
\end{longtable}

Table \ref{tbl:expected-series-lengths-alternative-ps} shows the new
\(P_E(i)\) when re-applying Equation \ref{eq:series-length} for each new
\(p\). I observe that \(P_E(i)\) is notably smaller for \(p_{1,2,3}\)
when \(i \in [10, 11]\) and higher for \(i \in [6, 7]\), more closely
matching \(P_O(i)\).
